---
title: "OAS Paths Analysis"
date: "<sup>Last updated: `r format(Sys.time(), '%Y-%m-%d')`</sup>"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
require('knitr')
knitr::opts_chunk$set(echo = FALSE)
source("db_connection.R", local = knitr::knit_global())
source("utils.R", local = knitr::knit_global())
library("ggplot2")
```

```{r, child=c('header.section.Rmd')}
```

<sup>[Back to summary](oas_summary.md) | [View related issues](https://github.com/postman-open-technologies/knowledge-base/labels/oas%3Apaths)</sup>

# Findings

## How is the path property used in APIs?
```{r oas_paths_summary, child=c('oas_paths_summary.section.Rmd')}
```

```{r, eval=TRUE }
query = "
select 
	bucket, 
	case 
		when bucket=0 then '0'
		when bucket=1 then '1-5'
		when bucket=2 then '6-10'
		when bucket=3 then '11-20'
		when bucket=4 then '21-50'
		when bucket=5 then '50+'
	end as label,
	count(*)
from (
select uri, count(*) as count , 
	width_bucket(count(*), array[1,6,11,21,51]) as bucket
	from (
	  select uri, jsonb_object_keys(json_data->'paths') as keys
	  from vw_oas_resource
	  where (json_meta->'isValid')::boolean
	) as t1 group by 1 
) as t2
group by 1,2 order by bucket
"
df = dbGetQuery(conn, query)
```

```{r oas_paths_buckets_barplot, include = TRUE, out.width="90%"}
df$label <- factor(df$label, levels = df$label)
ggplot(df, aes(x = label, y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.3) +
  ggtitle("Distribution of paths entries per API") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  xlab("Count of 'paths' entries") +
  ylab("Count of APIs")
```

## Which properties are used under paths?

```{r, eval=TRUE }
query = "
select jsonb_object_keys(value) as property, count(*) as n, count(*) / sum(count(*)) over() as pct
from vw_oas_resource_paths paths
where jsonb_typeof(value) = 'object'
group by 1 order by 2 desc
"
df = dbGetQuery(conn, query)
```

- Out of `r format_count(sum(df$n))` properties under paths, the most commonly used are ``r df$property[1]`` `r format_count_and_pct(df$n[1], df$pct[1])`, ``r df$property[2]`` `r format_count_and_pct(df$n[2], df$pct[2])`, ``r df$property[3]`` `r format_count_and_pct(df$n[3], df$pct[3])`, ``r df$property[4]`` `r format_count_and_pct(df$n[4], df$pct[4])`, and ``r df$property[5]`` `r format_count_and_pct(df$n[5], df$pct[5])`. 
- Beyond standard properties, numerous extensions were found. See table below for details.
- No significant variations were observed across specification versions (2.x vs 3.x) or collections
- Consult the [operations report](oas_paths_operations.md) for more in-depth analysis.q

<details>
<summary>Table: Counts and percentages of properties under paths</summary>
```{r, eval=TRUE, echo=FALSE, results='asis'}
kable(df)
```
</details>


# Methodology

Results are based on the content and count of entries in the `/paths` property in the APIs specifications.



