---
title: "OAS Servers Analysis"
date: "<sup>Last updated: `r format(Sys.time(), '%Y-%m-%d')`</sup>"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("db_connection.R", local = knitr::knit_global())
source("utils.R", local = knitr::knit_global())
```

```{r, child=c('header.section.Rmd')}
```
<sup>[View issues related to oas:servers](https://github.com/postman-open-technologies/knowledge-base/labels/oas%3Aservers)</sup>

# Overview
 
The `servers` property in OpenAPI provide connectivity information to target server(s) for the API. It replaces the `host`, `basePath` and `schemes` properties used in OpenAPI 2.
This differences between both versions add a bit to the complexity of the analysis. Note that more server entries may appears in the 3.x versions as the version 2.x was essentially limited to a single server entry (with multiple protocols). Version 3.x also introduced server variables which may obfuscate values for the host and other components of the URL.

# Findings

## How is the server property used in APIs?

```{r, child=c('oas_servers_summary.section.Rmd')}
```

## How often do APIs use localhost vs domain based servers? 

```{r, child=c('oas_servers_hosts.section.Rmd')}
```

*@TODO: compute API level statistics (localhost vs DNS vs mixed)*

# Methodology

Results are based on the content and count of entries in the `/servers` property in the OpenAPI 3.x specification, and its equivalent in version 2.0.

For the purpose of categorization across, we extract the host, protocol, and base name from all servers entries in 3.x to align on the 2.0 version

The host is then categorized as local if the name contains `localhost` or `127.0.0.1`, and  dns based if it is a valid domain name. This therefore excludes relative paths as well as entries that use variable templates.





