```{r host, eval=TRUE, echo=FALSE}
query = "
select count(*) as n_valid,
count(*) filter (where json_data->'host' is not null) as n_has_host,
count(*) filter (where json_data->'host' is not null) / count(*)::numeric as pct_has_host,
count(*) filter (where json_data->'basePath' is not null) as n_has_basepath,
count(*) filter (where json_data->'basePath' is not null) / count(*)::numeric as pct_has_basepath,
count(*) filter (where json_data->'schemes' is not null) as n_has_schemes,
count(*) filter (where json_data->'schemes' is not null) / count(*)::numeric as pct_has_schemes
from vw_oas_resource
where json_meta->'isValid' = 'true' and class='OpenAPI2'
"
df = dbGetQuery(conn, query)
```

- Out of `r format_count(df$n_valid[1])` valid OpenAPI2 specifications, `r format_count_and_pct(df$n_has_host[1], df$pct_has_host[1])` have a `host` property, 
`r format_count_and_pct(df$n_has_basepath[1], df$pct_has_basepath[1])` have a `basePath` property,
and `r format_count_and_pct(df$n_has_schemes[1], df$pct_has_schemes[1])` have a `schemes` property. 


```{r servers, eval=TRUE, echo=FALSE}
query = "
select count(*) as n_valid,
count(*) filter (where json_data->'servers' is not null) as n_has_servers,
count(*) filter (where json_data->'servers' is not null) / count(*)::numeric as pct_has_servers,
avg(jsonb_array_length(json_data->'servers')) filter (where json_data->'servers' is not null) as avg_servers,
min(jsonb_array_length(json_data->'servers')) filter (where json_data->'servers' is not null) as min_servers,
max(jsonb_array_length(json_data->'servers')) filter (where json_data->'servers' is not null) as max_servers
from vw_oas_resource
where json_meta->'isValid' = 'true' and class='OpenAPI3'
"
df = dbGetQuery(conn, query)
```

- Out of `r format_count(df$n_valid[1])` valid OpenAPI3 specifications, `r format_count_and_pct(df$n_has_servers[1], df$pct_has_servers[1])` have a `servers` property, with an average of `r format(round(df$avg_server[1],1))` server/entry, and a range of `r format_count(df$min_servers[1])`-`r format_count(df$max_servers[1])`. 
