---
title: "OAS Validity"
date: "<sup>Last updated: `r format(Sys.time(), '%Y-%m-%d')`</sup>"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
require('ggplot2')
knitr::opts_chunk$set(echo = FALSE)
source("db_connection.R", local = knitr::knit_global())
source("utils.R", local = knitr::knit_global())
```

```{r, child=c('header.section.Rmd')}
```

<sup>[Back to summary](oas_summary.md) | [View related issues](https://github.com/postman-open-technologies/knowledge-base/labels/oas%3Avalidity)</sup>

# Findings

## *How many valid OpenAPIs do we have?*
```{r validation, eval=TRUE, echo=FALSE}
query = "
select count(*) as n,
count(*) filter (where json_meta->'isValid' = 'true') as n_valid,
count(*) filter (where json_meta->'isValid' = 'true') / count(*)::numeric as pct_valid,
count(*) filter (where json_meta->'isValid' = 'false') as n_invalid,
count(*) filter (where json_meta->'isValid' = 'false') / count(*)::numeric as pct_invalid,
count(*) filter (where json_meta->>'class' = 'OpenAPI2') as n_openapi2,
count(*) filter (where json_meta->'isValid' = 'true' and json_meta->>'class' = 'OpenAPI2') as n_valid_openapi2,
count(*) filter (where json_meta->'isValid' = 'true' and json_meta->>'class' = 'OpenAPI2') / count(*) filter (where json_meta->>'class' = 'OpenAPI2')::numeric as pct_valid_openapi2,
count(*) filter (where json_meta->>'class' = 'OpenAPI3') as n_openapi3,
count(*) filter (where json_meta->'isValid' = 'true' and json_meta->>'class' = 'OpenAPI3') as n_valid_openapi3,
count(*) filter (where json_meta->'isValid' = 'true' and json_meta->>'class' = 'OpenAPI3') / count(*) filter (where json_meta->>'class' = 'OpenAPI3')::numeric as pct_valid_openapi3,
count(*) filter (where json_meta->'isValid' is null) as n_unknown
from vw_oas_resource
"
df = dbGetQuery(conn, query)
```

- Out of `r format_count(df$n[1])` entries, `r format_count(df$n_valid[1])` (`r format_pct(df$pct_valid[1])`%) are valid and `r format_count(df$n_invalid[1])` (`r format_pct(df$pct_invalid[1])`%) are invalid
- `r format_count(df$n_valid_openapi2[1])` out of the `r format_count(df$n_openapi2[1])` OpenAPI2 entries are valid (`r format_pct(df$pct_valid_openapi2[1])`%)
- `r format_count(df$n_valid_openapi3[1])` out of the `r format_count(df$n_openapi3[1])` OpenAPI3 entries are valid (`r format_pct(df$pct_valid_openapi3[1])`%)

```{r oas_validity_charts, echo=FALSE, include = TRUE}
category = c(rep('Valid',3), rep('Invalid',3))
version = rep( c("All", "OpenAPI2", "OpenAPI3"), 2)
percentage = c(df$pct_valid[1], df$pct_valid_openapi2[1], df$pct_valid_openapi3[1], 1-df$pct_valid[1], 1-df$pct_valid_openapi2[1], 1-df$pct_valid_openapi3[1])
data = data.frame(category, version, percentage)
ggplot(data, aes(fill=version, y=percentage, x=category)) + geom_bar(position="dodge", stat="identity")
```

## *Which top level paths of API specifications are the source of validation errors?*

<sup>See [source of errors](oas_validity_errors.md) for details and more in-depth analysis around validation issues.<sup>

```{r, child=c('oas_validity_errors_summary.section.Rmd')}
```

# Methodology
Results are based on the count of the `isValid` boolean property in the resource metadata. This flag is set by the `kb_oas_validation.py` script, which uses the JSON schemas published under the [OAI GitHub project](https://github.com/OAI/OpenAPI-Specification/tree/main/schemas). Validation is performed by the Python [jsonschema](https://github.com/python-jsonschema/jsonschema) package. Validation errors are saved in a local file for further analysis.
 