---
title: "OAS Validity: Source of Errors"
date: "<sup>Last updated: `r format(Sys.time(), '%Y-%m-%d')`</sup>"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
require('knitr')
require('ggplot2')
knitr::opts_chunk$set(echo = FALSE)
source("db_connection.R", local = knitr::knit_global())
source("utils.R", local = knitr::knit_global())
```

```{r, child=c('header.section.Rmd')}
```

<sup>[Back to validity](oas_validity.md) | [View related issues](https://github.com/postman-open-technologies/knowledge-base/labels/oas%3Avalidity)</sup>

# Findings

## *Which top level paths of API specifications are the source of validation errors?*

```{r, child=c('oas_validity_errors_summary.section.Rmd')}
```

<details>
<summary>Table: Count of errors by top property</summary>
```{r, eval=TRUE, echo=FALSE, results='asis'}
kable(df1)
```
</details>

<details>
<summary>Table: Count of APIs with 1 or more errors by top property</summary>
```{r, eval=TRUE, echo=FALSE, results='asis'}
kable(df2)
```
</details>

## *Which errors are found under `paths`?*

The majority of the API validation issues occur under the `path` property.

```{r, eval=TRUE, echo=FALSE}
query = "
select category, count(*) as n,  count(*) / sum(count(*)) over() as pct
from (
	select uri, repo_path, message, path, 
	case 
		when message ~ 'regexes'  then 'REGEX'
		when message ~ '''object'''  then 'NOTOBJ'
		when message ~ 'Unevaluated'  then 'UNEVAL'
		else 'OTH'
	end as category
	from  vw_oas_resource_validation_errors  
	where  path#>>'{0}' = 'paths' and jsonb_array_length(path) = 1
	order by category, message, repo_path
) as t1
group by 1
order by n desc
"
paths_df0 = dbGetQuery(conn, query)
```

 `r format_count(sum(paths_df0$n))` errors are found under at the first level (the properties right under `/paths`) ,which represents a small proportion of all `paths` errors (`r format_pct(sum(paths_df0$n) / df0$n_errors[1])`%), but are nonetheless important as they hide all underlying content.
 
They can be categorized as follows:

 - Regex mismatch: `r format_count_and_pct(paths_df0[paths_df0$category=='REGEX',]$n, paths_df0[paths_df0$category=='REGEX',]$pct)`. Raised when the property name does not start with a / or x-, essentially not matching the JSON schema  `'^/', '^x-'` regular expression. Note that the schema validation library aggregates all regex issues under a single error (not one per property). 
 - Not an object: `r format_count_and_pct(paths_df0[paths_df0$category=='NOTOBJ',]$n, paths_df0[paths_df0$category=='NOTOBJ',]$pct)`. Raised when the `paths` property is not a valid JSON {object}. Examples includes an empty [] or populated array, a string, a null value, etc.
 - Unevaluated properties: `r format_count_and_pct(paths_df0[paths_df0$category=='UNEVAL',]$n, paths_df0[paths_df0$category=='UNEVAL',]$pct)`. Raise when an invalid property name is used in a v3.1 specification.
 
 <details>
<summary>Table: Categorized count of errors under `paths` first level</summary>
```{r, eval=TRUE, echo=FALSE, results='asis'}
kable(paths_df0)
```
</details>

## *Which errors are found under `definitions` and `components`?*

*Analysis in progress*

## *Which errors are found at the root level?*

*Analysis in progress*


# Recommendations
- t.d.b.

# Methodology

Validation errors reported by the Python [jsonschema](https://github.com/python-jsonschema/jsonschema) package are stored in JSON format in the database table, with a view exposing each issue as a record. The validation metadata provides information on the JSON path where the error occured and an error message. To classify the nature of the issue, the error messages are matched to various categories using regular expressions.




